---
title: "P8106_stl2137_HW2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mgcv)
library(pdp)
library(earth)
library(caret)
```

The response variable is the out-of-state tuition (Outstate). 
```{r, results=FALSE, warning=FALSE}
school_dat <- read_csv("./College.csv") %>% 
  janitor::clean_names()

school_no_columbia_dat <- school_dat[-125,]
```

## Part A 

```{r}
predictors_df <- school_no_columbia_dat %>% 
  select(outstate, everything()) %>% 
  group_by(college) %>% 
  pivot_longer(
    apps:grad_rate, 
    names_to = "predictor",
    values_to = "values"
  )

predictors_df %>% 
  ggplot(
    aes(x = values, y = outstate, color = predictor)
  ) +
  geom_point(alpha = 0.25) +
  facet_wrap(. ~ predictor, ncol = 4)
```

## Part B
(b) Fit a smoothing spline model using Terminal as the only predictor of Outstate for a range of degrees of freedom, as well as the degree of freedom obtained by generalized cross- validation, and plot the resulting fits. Describe the results obtained.
```{r}
smooth_spline_fit <- smooth.spline(school_no_columbia_dat$terminal, school_no_columbia_dat$outstate)
smooth_spline_fit$df

terminal_lims <- range(school_no_columbia_dat$terminal)
terminal_grid <- seq(from = terminal_lims[1], to = terminal_lims[2])

pred_smooth <- predict(smooth_spline_fit, 
                       x = terminal_grid)

pred_sspline_df <- data.frame(pred = pred_smooth$y,
                             terminal = terminal_grid)

p <- ggplot(data = school_no_columbia_dat, 
            aes(
              x = terminal,
              y = outstate
            )) + geom_point() + theme_bw()

p + geom_line(
  aes(
    x = terminal, 
    y = pred), 
  data = pred_sspline_df,
  color = rgb(.8, .1, .1, 1)) + theme_bw()
```

## Part C
(c) Plot the results and explain your findings.
```{r}
gam_school_1 <- gam(outstate ~ apps + accept + enroll + top10perc + top25perc + f_undergrad + p_undergrad + room_board + books + personal + ph_d + s(terminal) + s_f_ratio + perc_alumni + expend + grad_rate, data = school_no_columbia_dat)

plot(gam_school_1)
```

## Part D

```{r}
# matrix of predictors 
school_no_columbia_dat <- school_no_columbia_dat[-1]
  
x <- model.matrix(outstate~.,school_no_columbia_dat)[,-1]
y <- school_no_columbia_dat$outstate

mars_grid <- expand.grid(degree = 1:2,
                         nprune = 2:10)

control1 <- trainControl(method = "cv", number = 10)

set.seed(1)
mars_fit <- train(x, y, 
                  method = "earth",
                  tuneGrid = mars_grid,
                  trControl = control1)

ggplot(mars_fit)

mars_fit$bestTune

coef(mars_fit$finalModel) 
```

```{r}
partial_school_1 <- partial(mars_fit, pred.var = c("f_undergrad"), grid.resolution = 10) %>% autoplot()

partial_school_2 <- partial(mars_fit, pred.var = c("f_undergrad", "ph_d"), grid.resolution = 10) %>% 
      plotPartial(levelplot = FALSE, zlab = "yhat", drape = TRUE, 
                  screen = list(z = 20, x = -60))

grid.arrange(partial_school_1, partial_school_2, ncol = 2)
```

## Part E

```{r}
### Grabbing Columbia observation
columbia_dat <- school_dat[125,]

columbia_gam <- predict(gam_school_1, newdata = columbia_dat)

columbia_mars <- as.numeric(predict(mars_fit, newdata = columbia_dat))

```

Based off the GAM model, we predict that the out-of-state tuition at Columbia University is `r columbia_gam`. Based off the MARS model, we predict that the out-of-state tuition at Columbia University is `r columbia_mars`. Between the two models, the GAM model predicts the out-of-state tuition for Columbia to be higher by `r columbia_gam - columbia_mars` compared to MARS model. 
